---
title: "dataset_description_assignment"
author: "Rose Graf"
date: "2/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries, set working Directory and load CSV
```{r}
library(pastecs)
#library(ggplot2)
library(devtools)
library(tidyverse)
library(gridExtra)
library(grid)
library(hrbrthemes)
library(ggpmisc)
setwd("/Users/rosegraf/Desktop/Bioinformatics/Bootcamp")
mydata <- read.csv("rose-dataset.csv")
```

Looking at the data
```{r}
head(mydata, 10)

```

Descriptive statistics of major variables
```{r}
summary(mydata)
options(scipen=100)
options(digits=1)
stat.desc(mydata, norm = TRUE)
```

Distribution of age of patients
```{r}
line_label <- paste("Ave Age", toString(round(mean(mydata$age), 1)), sep = " ")
qplot(mydata$age, binwidth = 5) + 
  xlab('Age') + 
  ylab('Count') + 
  labs(title = 'Age of Patients') + 
  geom_histogram(fill = '#557d82', binwidth = 5, color = 'grey') +
  geom_vline(xintercept = mean(mydata$age)) + 
  geom_text(aes(x = mean(mydata$age) + 5, label = line_label , y=max(mydata$age) + 13), colour="black", angle=0, vjust = 1.2, text=element_text(size=11))

```
box and whisker of male vs female height
```{r}
p1 <- ggplot(data = mydata, mapping =  aes(x = gender, y = height)) +
  geom_boxplot(outlier.colour = "#e8c76b") + 
  labs(title = "Height by Gender")
p1
```
box and whisker of bmi in male vs female
```{r}
p2 <- ggplot(data = mydata, mapping =  aes(x = gender, y = bmi)) +
  geom_boxplot(outlier.colour = "#e8c76b") + 
  labs(title = "BMI by Gender")
p2
```
```{r}
p3 <- ggplot(data = mydata, mapping =  aes(x = gender, y = systolic)) +
  geom_boxplot(outlier.colour = "#e8c76b") + 
  labs(title = "Systolic by Gender")
p3
```
```{r}
p4 <- ggplot(data = mydata, mapping =  aes(x = gender, y = diastolic)) +
  geom_boxplot(outlier.colour = "#e8c76b") + 
  labs(title = "Diastolic by Gender") +
  theme_set(theme_gray(base_size = 8 ))
p3
```
Combining Plots made
```{r}
title1=textGrob("Gender comparison across health statistics", gp=gpar(fontsize=16,font=8))
grid.arrange(p1, p2, p3, p4, nrow = 1, top = title1)

```


Conclusion: there is a correlation between bmi and systolic blood pressure, which is significant. 
```{r}
#create linear model and find out statistics and coefficients for bmi and systolic
linearModSys <- lm(systolic ~ bmi, mydata)
#print(linearModSys)
summary(linearModSys)

#making the linear regression for bmi and systolic
line_label2 <- paste("y = ", toString(round(linearModSys$coefficients[1], 2)), "+", toString(round(linearModSys$coefficients[2], 2)), "x", sep = " ") 
line_label2

#plot with linear trend (black) and confidence interval (green)
p7 <- ggplot(mydata, aes(x=bmi, y=systolic)) +
  geom_point(color="#3c7d3e") +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum() +
  annotate("text", x = 52, y = 190, label = line_label2)+ 
  labs(title = "Relationship of Systolic blood pressure and bmi")
p7 


```
Conclusion: there is a significant correlation between diastolic and bmi. 
```{r}
#create linear model and find out statistics and coefficients for bmi and systolic
linearModDia <- lm(diastolic ~ bmi, mydata)
#print(linearModDia)
summary(linearModDia)

#making the label for linear regression for bmi and diastolic
line_label2 <- paste("y = ", toString(round(linearModDia$coefficients[1], 2)), "+", toString(round(linearModDia$coefficients[2], 2)), "x", sep = " ") 

#plot of bmi and diastolic with linear trend (black) and confidence interval (green)
p8 <- ggplot(mydata, aes(x=bmi, y=diastolic)) +
  geom_point(color="#3c7d3e") +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum() +
  annotate("text", x = 52, y = 120, label = line_label2) + 
  labs(title = "Relationship of diastolic and bmi")
p8
```

Overall conclusion: bmi is positively correlated to blood pressure

